---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dataobservatory

<!-- badges: start -->
[![Project Status: Active. The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.5068765)](https://doi.org/10.5281/zenodo.5068765)
[![Codecov test coverage](https://codecov.io/gh/dataobservatory-eu/dataobservatory/branch/master/graph/badge.svg)](https://codecov.io/gh/dataobservatory-eu/dataobservatory?branch=master)
[![Follow author](https://img.shields.io/twitter/follow/antaldaniel.svg?style=social)](https://twitter.com/intent/follow?screen_name=antaldaniel)
[![Follow rOpenGov](https://img.shields.io/twitter/follow/ropengov.svg?style=social)](https://twitter.com/intent/follow?screen_name=ropengov)
<!-- badges: end -->

The goal of dataobservatory is to facilitate the automated documentation, and the automated recording of descriptive and administrative (statistical processing) metadata for datasets. It also helps recording information about the computational environment to increase reproducability.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("dataobservatory-eu/dataobservatory")
```
## The dataset Class

The `dataset` S3 class is an extension of the data frame and tibble class.  It has some important metadata attributes that facilitate the automated documentation of the dataset. Furthermore, it has an adequate `print` and `summary` method.

```{r example, echo=FALSE}
library(dataobservatory)
data("small_population")
small_population_dataset <- dataset (x= small_population,
                                     dataset_code = "small_population_total",
                                     dataset_title = "Population of Small European Countries",
                                     freq = "A",
                                     unit = "NR",
                                     unit_name = "number")
```



```{r print-dataset}
small_population_dataset
```

```{r summary-dataset}
summary(small_population_dataset)
```

```{r cite-dataset}
small_population_datacite <- datacite_dataset(
  dataset = small_population_dataset,
  Subject = "Demography",
  Creator = "Joe, Doe")
```


## Descriptive Metadata

The `datacite` class (see `?datacite()`) is a modification of a data frame (tibble) object, and it creates all the mandatory and recommended fields of the [DataCite](https://support.datacite.org/docs/datacite-metadata-schema-v44-properties-overview) metadata schema for a dataset.  It also covers all the properties in the more general Dublin Core standard, but in some cases, the property name is different (and follows the DataCite naming convention.)  

The descriptive metadata can be added with the `datacite()` constructor (see: `?datacite` ) or the `datacite_dataset()` helper function. or read the [DataCite Descriptive Metadata](http://r.dataobservatory.eu/articles/datacite.html) vignette article. 

The `datacite` class can automatically connected to many scientific repositories, including Zenodo.  In later versions, this will enable the user to upload the new created dataset (version) and receive a digital object identifier (version), or DOI(version) for the dataset.

See more about the metadata concepts applied in the [FAIR Data and the Added Value of Rich Metadata](https://contributors.dataobservatory.eu/FAIR-data.html) chatper of the [Automated Observatory Contributors’ Handbook](https://contributors.dataobservatory.eu/).

```{r print-datacite}
print(small_population_datacite)
```

## Administrative Metadata

The statistical processing information can be added with the not fully implemented `codebook` class. Read the [The codebook Class](http://r.dataobservatory.eu/articles/codebook.html) vignette article.

The `codebook` S3 class (not yet fully documented and does not have yet and independent constructor) records the statistical processing metadata of a dataset. 

It contains a full codebook following SDMX statistical metadata codelist standards, furthermore, it records the Session Information of all processing steps, and adds to the descriptive metadata the R packages or software code that generated the results.

For example, the annual observations follow the [SDMX Code List for Frequency 2.1 (CL_FREQ)](https://sdmx.org/?page_id=3215/)) definition, and they can be translated to the `ISO 8106` time metadata standard, too.

```{r}
add_frequency("A", "list")
```
```{r}
add_sessioninfo()
```

## Contributor Code of Conduct

Please note that the `dataobservatory` project is released with a [Contributor Code of Conduct](https://www.contributor-covenant.org/version/2/0/code_of_conduct/). By contributing to this project, you agree to abide by its terms.
