---
title: "The codebook Class"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The codebook Class}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=TRUE}
library(dataobservatory, quietly = TRUE)
data("small_population")
small_population_dataset <- dataset (x= small_population,
                                     dataset_code = "small_population_total",
                                     dataset_title = "Population of Small European Countries",
                                     freq = "A",
                                     unit = "NR",
                                     unit_name = "number")
```

The `codebook` S3 class (not yet fully documented and does not have yet and independent constructor) records the statistical processing metadata of a dataset. It as a print method.

It contains a full codebook following SDMX statistical metadata codelist standards, furthermore, it records the Session Information of all processing steps, and adds to the descriptive metadata the R packages or software code that generated the results.

```{r}
small_population_codebook <- codebook_dataset (small_population_dataset)
is.codebook(small_population_codebook)
small_population_codebook
```

## Frequency of Observation {#freq}

For example, the annual observations follow the [SDMX Code List for Frequency 2.1 (CL_FREQ)](https://sdmx.org/?page_id=3215/)) definition, and they can be translated to the `ISO 8106` time metadata standard, too.

```{r}
add_frequency("A", "list")
```
## Observation Status {#obs_status}

The `add_obs()` implements the [SDMX Code List for Observation Status 2.2 (CL_OBS_STATUS)](https://sdmx.org/?sdmx_news=new-version-of-code-list-for-observation-status-version-2-2) defintiion for observation status. 

Furthermore, with **E**stimated values and **I**mputed values, the software code and computational environment information can be recoded (with `add_related_items()`).

```{r}
add_obs("E", "list")
```

## Unit information {#unit}

Only Eurostat uses hundreds of codelists, most of them for various measurement information. Currently the unit information is not validated in the package. However, this is not a problem when data is imported from other statistical services, because they apply appropriate unit information.

```{r unit-info}
add_unit("M_EUR", "Million euros", 
         admin_format = 'list')
```
## Consolidated Coodbook

```{r}
consolidated_codebook <- codebook()
is.codebook(consolidated_codebook)
print(consolidated_codebook)
```

