---
title: "DataCite Descriptive Metadata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DataCite Descriptive Metadata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `datacite` class (see `?datacite()`) is a modification of a data frame (tibble) object, and it creates all the mandatory and recommended fields of the [DataCite](https://support.datacite.org/docs/datacite-metadata-schema-v44-properties-overview) metadata schema for a dataset.  It also covers all the properties in the more general Dublin Core standard, but in some cases, the property name is different (and follows the DataCite naming convention.)  

```{r datacite-properties, message=FALSE, warning=FALSE}
library(dataobservatory, quietly = TRUE)
data("datacite_properties")
datacite_properties
```
The `datacite` class is equipped with a construction and helper functions that create the more complex metadata properties as JSON (default) or list objects. Not all the helper functions are fully validated, but a fully datacite object can be easily obtained. 

```{r setup, message=FALSE, warning=FALSE}
library(jsonlite, quietly = TRUE)
data("small_population")
small_population_dataset <- dataset (
  x= small_population,
  dataset_code = "small_population_total",
  dataset_title = "Population of Small European Countries",
  freq = "A",
  unit = "NR",
  unit_name = "number")

attributes (small_population_dataset)
```

The `datacite_dataset()` function fills out the `datacite` object with default properties. There are some properties where no default is given, these properties are also recorded in the `dataset` class as attributes.

```{r cite}
small_population_datacite <- datacite_dataset(
  dataset = small_population_dataset,
  keywords = c("greendeal", "Demography", "Testing"),
  description = "Example dataset with three small countries",
  Subject = "Demography",
  Creator = "Joe, Doe")
```

```{r}
small_population_dataset
```

```{r}
small_population_datacite
```
```{r}
is.datacite(small_population_datacite)
```

The `Description` property has three mandatory elements:

- The `Abstract` is a short, textual description.
- In the `TechnicalInfo` sub-field, we record automatically the `utils::sessionInfo()` for computational reproducability.
- In the `Other` sub-field, we record the keywords for structuring the observatory.

```{r}
small_population_description <- small_population_datacite$Description
jsonlite::fromJSON(small_population_description)
```
## Complex Properties

The Creator property may have affiliation data, and ORCiD identifier present, By default, we create a json representation:

```{r}
jane <- add_creators("Jane", "Doe", 
                     affiliation = "Green Deal Data Observatory", 
                     orcid = "0000-0000-0000-0000")
jane
```


## Working with lists

There are many Properties which can form a list.  For example, there may be several related items present.

```{r, eval=FALSE}
relitems <- add_related_items(
  RelatedItem = "Population on 1 January by age and sex", 
  relatedItemIdentifier = add_identifiers(
    id = "demo_pjan", 
    relatedItemType = "Dataset",
    relatedItemIdentifierType = "URL",
    URI = "https://appsso.eurostat.ec.europa.eu/nui/show.do?dataset=demo_pjan/"
  ))

relitems
```
```{r, eval=FALSE}
package_citation <- citation("dataobservatory")
relitems <- add_related_items(
  RelatedItem = package_citation$title, 
  relatedItemIdentifier = add_identifiers(
    id = "dataobservatory", 
    Version = package_citation$note),
    related_items = relitems
  )

relitems
```






```{r}
print(small_population_datacite)
```

```{r, eval=FALSE}
small_population_datacite$RelatedItem <- create_json_text(as.list(relitems))
```


```{r, eval=FALSE}
small_population_datacite
```

