---
title: "The dataset Class"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The dataset Class}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(dataobservatory, quietly = TRUE)
```

The `dataset` S3 class is an extension of the data frame and tibble class.  It has some important metadata attributes that help the documentation of the dataset. Furthermore, it has an adequate `print` and `summary` method.

```{r small-population, echo=FALSE, message=FALSE, warning=FALSE}
data("small_population")
small_population_dataset <- dataset (x= small_population,
                                     dataset_code = "small_population_total",
                                     dataset_title = "Population of Small European Countries",
                                     freq = "A",
                                     unit = "NR",
                                     unit_name = "number")

attributes (small_population_dataset)
```
```{r}
print(small_population_dataset)
```
```{r}
summary(small_population_dataset)
```
```{r}
is.dataset(small_population_dataset)
```
## Adding Metadata

The descriptive metadata can be added with the `datacite()` constructor (see: `?datacite` ) or the `datacite_dataset()` helper function. or read the [DataCite Descriptive Metadata](http://r.dataobservatory.eu/articles/datacite.html) vignette article.

The statistical processing information can be added with the not fully implemented `codebook` class. Read the [The codebook Class](http://r.dataobservatory.eu/articles/codebook.html) vignette article.

